---
title: "My Awesome Report"
author: "Claudiu Forgaci"
output: html_document
---

```{r, include=FALSE}
library(tidyverse)
# Make sure that gapminderDataFiveYear.tsv is in the same directory with this Rmd file 
gapMinder <- read.delim("gapminderDataFiveYear.tsv") 
```

# Heading

## Sub-heading

Text...

### Sub-sub-heading

Text...

## Introduction

This report uses the **tidyverse** package along with the *gapminder* dataset, which has columns that include:

1.  country

-   Name of country

2.  year

-   Year of observation

3.  pop

-   Population of country

4.  continent

-   Continent where the country is located

5.  lifeExp

-   Life expectancy

6.  gdpPercap

-   GDP per capita

## Insert table

The average GDP per capita grouped by `continent`.

```{r}
gapMinder %>%   
  # filter(!is.na(continent)) %>%   
  group_by(continent) %>%   
  summarise(mean_gdpPercap = mean(gdpPercap)) %>%   
  knitr::kable(caption = "We can also add a caption.",
               col.names = c("Continent", "Mean GDP per Capita"))
```

## Inline R code

```{r}
# Replace the value of cont with your continent of choice
cont <- "Europe"
mean_continent <- gapMinder |>  
  group_by(continent) |>  
  summarise(mean_lifeExp = mean(lifeExp)) |> 
  filter(continent == cont)
mean_continent
```

The average life expectancy in `r cont` is `r round(mean_continent$mean_lifeExp, 2)`.

## Plots

```{r}
continents <- factor(gapMinder$continent)

plot(continents)
```
```{r echo=FALSE}
ggplot(data = gapMinder,
       mapping = aes(x = continent)) +
  geom_bar()
```

